\section{Expansion Constant Properties}

I talked to someone who wasn't convinced that all of the following useful lemmas
were true, so I went out and proved them.  Now I've typeset them.

\begin{lemma}
\label{lem:ec_rec}
Given an expansion constant $c$ as defined in Definition \ref{def:int_dim}, for any
ball $B_S(p, 2^i \Delta)$ where $p \in S$, $i \in \mathbb{N}$, $\Delta > 0$,

\begin{equation}
| B_S(p, 2^i \Delta) | \le c^i | B_S(p, \Delta) |.
\end{equation}
\end{lemma}

\begin{proof}
This follows from $i$ recursive applications of the definition of the expansion
constant.
\end{proof}

\begin{lemma}
\label{lem:ec_re}
Given an expansion constant $c$ as defined in Definition \ref{def:int_dim}, for any
ball $B_S(p, \alpha \Delta)$ where $p \in S$, $\alpha \in \Real$, $\alpha \ge 1$,
$\Delta > 0$,

\begin{equation}
| B_S(p, \alpha \Delta) | \le c^{\log_2 \alpha} | B_S(p, \Delta) |.
\end{equation}
\end{lemma}

\begin{proof}
By Definition \ref{def:int_dim} and Lemma \ref{lem:ec_rec}, it is obvious that
the statement holds when $\alpha \in \mathbb{N}$.  However, the definition of
expansion constant makes no particular bound on the number of points added when
the ball is not doubled but simply increased by some real factor.  We will
perform this proof by contradiction.

Assume that there exists some $\alpha \in \mathbb{R}$ where $\alpha > 1$ and

\begin{equation}
| B_S(p, \alpha \Delta) | > c^{\log_2 \alpha} | B_S(p, \Delta) |.
\end{equation}

By Definition \ref{def:int_dim}, we also know that

\begin{equation}
| B_S(p, \alpha \Delta) | \le c | B_S(p, (\alpha / 2) \Delta) |.
\end{equation}

Combining these two equations gives that

\begin{equation}
c | B_S(p, (\alpha / 2) \Delta) | > c^{\log_2 \alpha} | B_S(p, \Delta) |
\end{equation}

\noindent or alternately

\begin{equation}
c^{(1 - \log_2 \alpha)} | B_S(p, (\alpha / 2) \Delta) | > | B_S(p, \Delta) |.
\end{equation}

First, assume that $\alpha < 2$.  Then, $|B_S(p, (\alpha / 2) \Delta)| \le
|B_S(p, \Delta)|$.  Thus,

\begin{equation}
c^{(1 - \log_2 \alpha)} | B_S(p, \Delta) | > | B_S(p, \Delta) |
\end{equation}

\noindent which is a contradiction because $c^{(1 - \log_2 \alpha)} > 1$.

Now, assume that $\alpha > 2$.  Define $\beta = 2^i \alpha$ where $i \in
\mathbb{N}$ is such that $1 < \beta < 2$.  Then,

\begin{equation}
| B_S(p, (2^i \beta / 2) \Delta) | \le c^i | B_S(p, (\beta / 2) \Delta) |.
\end{equation}

Because $\beta < 2$, $|B_S(p, (\beta / 2) \Delta)| \le |B_S(p, \Delta)|$.  Thus,

\begin{equation}
|B_S(p, (2^i\beta / 2) \Delta)| \le c^i | B_S(p, \Delta) |.
\end{equation}

From earlier, this implies that

\begin{equation}
\label{eqn:contr}
(c^{(1 - \log_2 (2^i \beta))}) (c^i) |B_S(p, \Delta) | > |B_S(p, \Delta) |.
\end{equation}

Note that $(c^{(1 - \log_2 (2^i \beta))}) (c^i) = c^{(1 - \log_2 \beta)}$, which
is greater than $1$ because $\beta < 2$.  This means that Equation
\ref{eqn:contr} is a contradiction, and thus, for all $\alpha \in \mathbb{R}$,
the statement of the lemma holds.
\end{proof}

\begin{lemma}
\label{lem:ec_neg}
Given an expansion constant $c$ as defined in Definition \ref{def:int_dim}, for
any ball $B_S(p, 2^{-i} \Delta)$ where $p \in S$, $i \in \mathbb{N}$, $\delta >
0$,

\begin{equation}
|B_S(p, 2^{-i} \Delta)| \ge c^{-i} |B_S(p, \Delta)|.
\end{equation}
\end{lemma}

\begin{proof}
Observe that $|B_S(p, \Delta)| \le c^i |B_S(p, 2^{-i} \Delta)|$.  Simple
algebraic manipulation yields the result.
\end{proof}

\begin{lemma}
Given an expansion constant $c$ as defined in Definition \ref{def:int_dim}, for
any ball $B_S(p, \alpha \Delta)$ where $p \in S$, $\alpha \in \mathbb{R}$,
$0 < \alpha \ge 1$, $\delta > 0$,

\begin{equation}
|B_S(p, \alpha \Delta)| \ge c^{\log_2 \alpha} |B_S(p, \Delta)|.
\end{equation}
\end{lemma}

\begin{proof}
(idea) This is a relatively straightforward combination of the proofs of
Lemmas \ref{lem:ec_neg} and \ref{lem:ec_re}.
\end{proof}

Two important observations:

\begin{itemize}
  \item We can place an upper bound on the number of points added as a ball
grows an arbitrary amount.

  \item We can place a lower bound on the number of points removed as a ball
shrinks an arbitrary amount.
\end{itemize}
