\documentclass[twoside,11pt]{article} % For LaTeX2e
\usepackage{jmlr2e}
\usepackage{hyperref}
\hypersetup{colorlinks,
            urlcolor=blue,
            citecolor=green,
            linkcolor=red}
\usepackage{url}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09
\usepackage{mathrsfs}
%\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{color}
\usepackage{mathtools}

\usepackage{pstricks}
\usepackage{tikz}
\usepackage{subcaption}

\begin{document}

\noindent Dear Drs. Murphy and Schölkopf: \\

\medskip

My co-authors (William March, Parikshit Ram, and Dongryeol Lee) would like to
submit our manuscript "Plug-and-play dual-tree algorithm runtime analysis" for
publication in the Journal of Machine Learning Research.  This paper establishes
a generalized technique for proving runtime bounds on the class of algorithms
known as dual-tree algorithms.  These algorithms are commonly used for
computations such as nearest neighbor search, range search, kernel density
estimation, and other applications.  For each of the three problems just listed,
we have shown linear runtime bounds using our generalized bounds (though it must
be noted that these bounds depend on the intrinsic dimensionality of the
dataset). \\

\medskip

Each co-author is aware of this submission and consent to its review, and the
paper has not been submitted to any other venues.  There are no conflicts of
interest, since no co-author has recently collaborated with any action editor.
\\
\medskip

Our suggestions for action editors are those who have published papers relating
to dual-tree algorithms in the past or are familiar with the approach and
existing theory.  These include Sanjoy Dasgupta, Nando de Freitas, and possibly
Guy Lebanon. \\
\medskip

If there is anything we have neglected to do in the submission process, please
let us know and we will rectify the issue immediately. \\

\medskip

\noindent Thank you for your consideration, \\

\medskip

\noindent Ryan Curtin

\end{document}
